---
title: "STAT 201: Midterm 1 Practice 2"
format: pdf
---

```{r message = F, echo = F}
# load your libraries here for wrangling, plotting, and making pretty tables
library(readr)
library(tidyverse)
library(kableExtra)
```

A Pell Grant is a need-based federal grant for undergraduate students paying for college. Students are automatically considered for a Pell Grant award when they submit a FAFSA.

Each case in the dataset `pell` represents the Pell Grant award to a given school in a given year.

-   `state`: state/territory shortcode
-   `award`: total award amount in USD
-   `recipient`: total number of recipients
-   `name`: name of college/university
-   `session`: meeting ID
-   `year`: year

We also have a dataset called `states` that provide information on the states and territories of the US. The variables are:

-   `NAME`: name of the state/territory
-   `Abbreviation`: state/territory shortcode
-   `Type`: political devision ("state" or "territory")

Run the following code chunk to load in the two datasets:

```{r echo = F}
pell <- read_csv("https://raw.githubusercontent.com/midd-stat201-spring2025/midd-stat201-spring2025.github.io/refs/heads/main/data/pell.csv", show_col_types = F)
states <- read_csv("https://raw.githubusercontent.com/midd-stat201-spring2025/midd-stat201-spring2025.github.io/refs/heads/main/data/states.csv", show_col_types = F)
```

### Exercise 1

Let's clean and wrangle the data a bit.

-   Remove any case where at least one of the following is true:
    -   The school is missing information about the award amount
    -   The school had less than 1 student receiving a Pell Grant
    -   The school receive an award of 0 dollars but had more than 0 students recieve a Pell grant
-   Create a new variable called `award_pp` that represents the award amount per person for each school
-   Create a new variable called `decade` that takes the value:
    -   "1990s" if the award was granted in the 1990s
    -   "2000s" if the award was granted in the 2000s
    -   "2010s" if the award was granted in the 2010s

Store your cleaned and wrangled data frame back into `pell`.

```{r echo = F}
pell <- pell |>
  filter(!is.na(award)) |>
  filter(!(award == 0 & recipient > 0)) |>
  filter(recipient > 0)
pell <- pell |>
  mutate(decade = case_when(
    year < 2000 ~ "1990s",
    year >= 2000 & year < 2010 ~ "2000s",
    T ~ "2010s"
  )) |>
  mutate(award_pp = award/recipient)
```

### Exercise 2

Make a visualization the shows the distribution of the award amount per person for each decade. Have informative labels and titles. Interpret what you see.

```{r echo = F, eval = F}
pell |>
  ggplot(aes(x  = decade, y = award_pp)) +
  geom_boxplot() 
```

**Answer**:

### Exercise 3

Now let's bring in the information about states vs territories. Combine the two datasets into a single dataset called `pell2` that retains all observations about pell grants awards for which we also have information about the type of political division (i.e. state or territory).

```{r echo = F, eval = F}
pell2 <- pell |>
  inner_join(states, by = c("state" = "Abbreviation"))
```

### Exercise 4

For awards made in 2015, create a beautiful table that displays the mean and standard deviation of the award per person for each of the two political divisions. Then answer the question: did the two political divisions differ in the Pell grant awarded?

```{r echo = F, eval = F}
pell2 |>
  filter(year == 2015) |>
  group_by(Type) |>
  summarise(avg_award_pp = mean(award_pp), sd_award_pp = sd(award_pp)) |>
  kable()
```

**Answer**:

### Exercise 5

Display a beautiful table of the five U.S. states that received the largest statewide average award per person in the 2010s.

```{r echo = F, eval = F}
pell2 |>
  filter(Type == "State", decade == "2010s") |>
  group_by(Name)  |>
  summarise(state_avg = mean(award_pp)) |>
  ungroup() |>
  arrange(-state_avg) |>
  slice(1:5)
```

### Exercise 6

Re-create the following plot:

```{r echo = F}
pell |>
  filter(state == "VT", recipient < 500) |>
  mutate(college = if_else(name == "Middlebury College", "Middlebury", "Other")) |>
  ggplot(aes(x = recipient, y = award, col = decade, shape = college)) +
  geom_point() +
  labs(x = "Number of recipients", y = "Award amount (USD)",
       title = "Pell grant award by institution and year",
       subtitle = "Vermont institutions with less than 500 recipients")
```
